Local Spectral Clustering of Density Upper Level Sets

### 2019-07-03 neurips/local_spectral_density_clustering_arxiv.tex
completed reorganization of theorem 4

%%%%%% Begin Alden
\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\newcommand{\diam}{\rho}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\left\{#1\right\}_{n \in \mathbb{N}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\cut}{\mathrm{cut}}
% \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\Phibf}{\Phi_{u}}
\newcommand{\Psibf}{\Psi_{u}}
\newcommand{\taubf}{\tau_{u}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\TV}{\mathrm{TV}}

@@

Denote $\widetilde{G}_{n,r} = G_{n,r}[\Csig[\Xbf]]$; in general, we will use tilde notation to refer to quantities computed over $\Csig$ or over the induced subgraph $\widetilde{G}_{n,r}$. The goal of Theorem \ref{thm: mixing_time_upper_bound} is to show that, for universal constants $c_2, c_3,c_4 > 0$, with probability one
\begin{equation*}
\limsup_{n \to \infty} \tau_{\infty}(\widetilde{G}_{n,r}) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{c_4 \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
\end{equation*}
We follow a two-step approach typically used to establish upper bounds on the mixing time of Markov chains.

In the first step, we relate the mixing time $\tau_{\infty}(G)$ of a generic undirected graph $G = (V,E)$ to expansion properties of subsets $U \subseteq V$. We now build to a formal definition of these expansion properties. First, we recall the the \emph{cut} and \emph{volume} functionals over a graph, and introduce the \emph{degree} functional as well. For $u \in V$, $S \subseteq V$, 
\begin{equation*}
\cut(S;G) = \sum_{u \in S} \sum_{v \in S^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(S;G) = \sum_{u \in S} \deg(u;G).
\end{equation*}
Additionally, we recall the \emph{normalized cut} $\Phi(S;G)$, defined (as in \eqref{eqn: normalized_cut}) as
\begin{equation*}
\Phi(S;G) = \frac{\cut(S; G)}{\min\set{\vol(S; G),\vol(S^c; G)}}.
\end{equation*}

@@

The following Proposition establishes an upper bound on the mixing time $\tau_{\infty}(G)$ in terms of the local spread $s(G)$ and conductance $\Phi(G)$.
\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in v} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}

The second step in our approach is to lower bound the local spread and conductance over the particular graph $\widetilde{G}_{n,r}$. The following result provides the desired lower bounds, which hold asymptotically as $n \to \infty$,
\begin{proposition}
	\label{prop: local_spread_conductance}
	Under the setup and conditions of Theorem \ref{thm: mixing_time_upper_bound}, with probability one,
	\begin{equation}
	\label{eqn: local_spread}
	\liminf_{n \to \infty} s(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 \nu_d r^d}{20\Lambda_{\sigma}} 
	\end{equation}
	and
	\begin{equation}
	\label{eqn: conductance}
	\liminf_{n \to \infty} \Phi(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}}.
	\end{equation}
\end{proposition}

Taking these results as given, the proof of the theorem is more or less complete. We note that since $r$ is fixed, \textcolor{red}{with probability one} $\liminf_{n \to \infty} \min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}_{n,r}) = \infty > 10$, and we may therefore apply Proposition \ref{prop: pointwise_mixing_time} with respect to $\widetilde{G}_{n,r}$. Along with Lemma \ref{prop: local_spread_conductance}, this yields
\begin{equation*}
\tau_{\infty}(\widetilde{G}_{n,r}) \leq \frac{2^{27}\Lambda_{\sigma}^4\rho^2L^2 d^3}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{28800\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3 \log^2\left(\frac{280\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3
\end{equation*}
The first term dominates the middle term, and the theorem follows after an appropriate choice of $c_2$ and $c_3$.

@@

We first generalize some notation from the main text. Let $A$ be the adjacency matrix of an undirected graph $G = (V,E)$, and $D$ the associated diagonal degree matrix. The lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf{} + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$.  Denote the $t$-step probability distribution of this random walk, originating from a particular $v \in V$, as $q^{(t)}: V \times V \to [0,1]$, $q^{(t)}(v,u) = e_v \Wbf^t e_u$.

Consider the uniform distance\footnote{Note $d_{\textrm{unif}}$ is not formally a distance as it is not symmetric.} between $q_v^{(t)} := q^{(t)}(v,\cdot)$ and $\pi$, given by
\begin{equation*}
d_{\textrm{unif}}(q_v^{(t)},\pi) = \max_{u \in V} \set{\frac{\pi(u) - q_v^{(t)}(u)}{\pi(u)}}.
\end{equation*}
Our goal is to demonstrate that for sufficiently large $t$, $d_{\textrm{unif}}(q_v^{(t)},\pi) \leq \frac{1}{4}$ for every $v \in V$. To do so, we introduce the \emph{total variation distance} between the distributions $q_v^{(t)}$ and $\pi$, given by
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sum_{u \in V} \abs{q_v^{(t)}(u) - \pi(u)}.
\end{equation*}
As we will see, an upper bound on the uniform distance can be obtained from an analogous upper bound on the total variation distance. First, however, we upper bound the total variation distance $||q_v^{(t)} - \pi||_{\mathrm{TV}}$ as a function of the local spread $s(G)$, the conductance $\Phi(G)$ and the number of steps $t$.
\begin{lemma}
	\label{lem: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1$,
	\begin{align*}
	\norm{q_v^{(t + 3)} - \pi}_{\mathrm{TV}} \leq \max\set{as(G), \frac{1}{8} + \frac{1}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} 
	\\
	~~~~~~~~~~~~~~~~~~~~~ + \left(\frac{1}{1 - 2as(G)} +  \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t
	\end{align*}
\end{lemma}

As mentioned, a bound on the total variation distance implies a corresponding bound on the uniform distance $d_{\mathrm{unif}}$, given by the following result:
\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{14} \max \set{1, \frac{1}{s(G)}}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}

\noindent Taking these Lemmas as given, we proceed that show that for
\begin{equation*}
\tau_1 = \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
\end{equation*}
the uniform distance $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$, which proves the claim of Proposition \ref{prop: pointwise_mixing_time}. Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$. Then,
\begin{equation*}
\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}
\end{equation*}
and so by Lemma \ref{lem: tv_mixing_time},
\begin{equation*}
\norm{q_v^{(\tau_0 + 3)} - \pi}_{\mathrm{TV}} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{2}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
\end{equation*}
It is a well known fact \cite{montenegro2002} that if $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{4}$, then for any $0 < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{\mathrm{TV}} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
\begin{equation*}
\norm{q_v^{(\tau_1)} - \pi}_{\mathrm{TV}} \leq \frac{1}{14 s(G)}
\end{equation*}
and so by Lemma \ref{lem: tv_to_uniform_distance}, $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. The proof is complete once we have proved Lemmas \ref{lem: tv_mixing_time} and \ref{lem: tv_to_uniform_distance}.

@@

We will find it useful to introduce the \emph{Lovasz-Simonovits curve}, defined as
$h_z^{(t)}, t \geq 0$, 
\begin{equation*}
h_q^{(t)}(x) = \max_{w \in \mathcal{W}_x} \left\{ \sum_{u \in V} \left(q_z^{(t)}(u) - \pi(u)\right)w(u)\right\}
\end{equation*}
where the maximum is over:
\begin{align*}
\mathcal{W}_x = \left\{w: V \to [0,1] \Big\vert 0 \leq w(u) \leq 1~~\forall~u,~~\text{and}~~\sum_{u \in V} w(u) \pi(u) = x\right\}.
\end{align*}
The utility of the Lovasz-Simonovits curve is demonstrated by the following observations. First, note that
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sup_{S \subseteq V} \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation*}
where we use the standard notation $\pi(S) := \sum_{u \in S} \pi(u)$, and likewise for $q_v^{(t)}(S)$. Moreover, observe that for any $S \subseteq V$ and any integer $t \geq 0$,
\begin{equation}
\label{eqn: lovasz_simonivits_1}
\max \set{h_v^{(t)}(\pi(S)), h_v^{(t)}(1 - \pi(S))} \geq \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation}
Therefore, taking the maximum on both sides of \eqref{eqn: lovasz_simonivits_1}  over all $S \subseteq V$, we have
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \max_{0 \leq x \leq 1} h_v^{(t)}(x).
\end{equation*}
Now, take $z = e_vW^3$. Observe that $h_v^{(t+3)}(x) = h_z^{(t)}(x)$. To prove Lemma \ref{lem: tv_mixing_time}, it is therefore sufficient to show the desired upper bound holds with respect to $h_z^{(t)}(x)$ for all $x \in [0,1]$.

To show this upper bound, we split the interval $[0,1]$ into the subinterval $[\mu, 1 - \mu]$ and the remainder $[0,\mu) \cup (1 - \mu, 1]$, where $\mu = as(G)$. Let $\ell_{\mu}(x)$ be the linear interpolator between $h_z^{(0)}(\mu)$ and $h_z^{(0)}(1 - \mu)$, 
\begin{equation*}
\ell_{\mu}(x) = \frac{1 - \mu - x}{1 - 2\mu} h_z^{(0)}(\mu) + \frac{x - \mu}{1 - 2\mu}h_z^{(0)}(1 - \mu).
\end{equation*}

The following technical Lemma collects the upper bounds which hold over $[\mu, 1 - \mu]$ and $[0,\mu) \cup (1 - \mu, 1]$, respectively:
\begin{lemma}
	\label{lem: mixing_all_sets}
	For any $\mu \leq x \leq 1 - \mu$, 
	\begin{equation}
	\label{eqn: mixing_large_sets}
	h_z^{(t)}(x) \leq \ell_{\mu}(x) + \left(\frac{1}{1 - 2\mu} + \frac{1}{\mu}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t.
	\end{equation}
	For any $0 < x \leq \mu$ or $1 - \mu \leq x < 1$
	\begin{equation}
	\label{eqn: mixing_small_sets}
	h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{2^t} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
	\end{equation}
\end{lemma}
\noindent Taking Lemma \ref{lem: mixing_all_sets} as given, we have nearly completed our proof of Lemma \ref{lem: tv_mixing_time}. Note that for any $\mu \leq x \leq 1 - \mu$,
\begin{align*}
\ell_{\mu}(x) & \leq \max\set{h_z^{(t)}(\mu),h_z^{(t)}(1 - \mu)} \\
& \leq \max\set{as(G), \frac{1}{2^t} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
\end{align*}
with the latter inequality following from \eqref{eqn: mixing_small_sets}. Therefore, for any $x \in [0,1]$,
\begin{equation*}
h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{2^t} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} } + \left(\frac{1}{1 - 2as(G)} + \frac{1}{as(G)}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t
\end{equation*}
which is the desired result.

@@

We first prove \eqref{eqn: local_spread} and then \eqref{eqn: conductance}. First, we introduce some notation. Recall that $\widetilde{G}_{n,r} = G_{n,r}\bigl[\Csig[\Xbf]\bigr]$ denotes the subgraph induced by $\Csig[\Xbf]$. For $S \subseteq \Csig[\Xbf]$ and $u \in \Csig[\Xbf]$, we will write
\begin{equation*}
\widetilde{\cut}_{n,r}(S) := \cut(S; \widetilde{G}_{n,r}), \quad \widetilde{\deg}_{n,r}(u) := \deg(u; \widetilde{G}_{n,r}), \quad 
\widetilde{\vol}_{n,r}(S) := \vol(S; \widetilde{G}_{n,r})
\end{equation*}
and let $\widetilde{\pi}_{n,r}(u) = \widetilde{\deg}_{n,r}(u)/\widetilde{\vol}_{n,r}(\Csig[\Xbf])$ be the stationary distribution of the lazy random walk over $\widetilde{G}_{n,r}$. We also let $\widetilde{\Phi}_{n,r}(S) := \Phi(S; \widetilde{G}_{n,r})$ denote the normalized cut functional over $\widetilde{G}_{n,r}$.

>>>>

%%%%%% Begin Alden

\newcommand{\diam}{\rho}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\left\{#1\right\}_{n \in \mathbb{N}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\cut}{\mathrm{cut}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\Phibf}{\Phi_{u}}
\newcommand{\Psibf}{\Psi_{u}}
\newcommand{\taubf}{\tau_{u}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\TV}{\mathrm{TV}}

@@

Let $\widetilde{G}_{n,r} := G_{n,r}[\Csig[\Xbf]]$; in general, we will use tilde notation to refer to quantities computed over $\Csig$ or over the induced subgraph $\widetilde{G}_{n,r}$. The goal of Theorem \ref{thm: mixing_time_upper_bound} is to show that, for universal constants $c_2, c_3,c_4 > 0$, with probability one
\begin{equation*}
\limsup_{n \to \infty} \tau_{\infty}(\widetilde{G}_{n,r}) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{c_4 \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
\end{equation*}
We follow a two-step approach typically used to establish upper bounds on the mixing time of Markov chains.

In the first step, we relate the mixing time $\tau_{\infty}(G)$ of an arbitrary undirected graph $G = (V,E)$ to expansion properties of subsets $U \subseteq V$. We now build to a formal definition of these expansion properties. First, we recall the the \emph{cut} and \emph{volume} functionals over a graph, and introduce the \emph{degree} functional as well. For $u \in V$, $S \subseteq V$, and $S^c = V \setminus S$,
\begin{equation*}
\cut(S;G) = \sum_{u \in S} \sum_{v \in S^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(S;G) = \sum_{u \in S} \deg(u;G).
\end{equation*}
Additionally, we recall the \emph{normalized cut} $\Phi(S;G)$, defined (as in \eqref{eqn: normalized_cut}) as
\begin{equation*}
\Phi(S;G) = \frac{\cut(S; G)}{\min\set{\vol(S; G),\vol(S^c; G)}}.
\end{equation*}

@@

The following Proposition establishes an upper bound on the mixing time $\tau_{\infty}(G)$ in terms of the local spread $s(G)$ and conductance $\Phi(G)$.
\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in V} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}

The second step in our approach is to lower bound the local spread and conductance over the neighborhood graph $\widetilde{G}_{n,r}$. In the following result we give lower bounds for both quantities, which hold asymptotically as $n \to \infty$.
\begin{proposition}
	\label{prop: local_spread_conductance}
	Under the setup and conditions of Theorem \ref{thm: mixing_time_upper_bound}, with probability one,
	\begin{equation}
	\label{eqn: local_spread}
	\liminf_{n \to \infty} s(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 \nu_d r^d}{20\Lambda_{\sigma}} 
	\end{equation}
	and
	\begin{equation}
	\label{eqn: conductance}
	\liminf_{n \to \infty} \Phi(\widetilde{G}_{n,r}) \geq \frac{\lambda_{\sigma}^2 r}{\Lambda_{\sigma}^2 2^{13} \rho L \sqrt{d}}.
	\end{equation}
\end{proposition}

Taking these results as given, the proof of Theorem \ref{thm: mixing_time_upper_bound} is more or less complete. We note that since $r$ is fixed, with probability one $\liminf_{n \to \infty} \min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}_{n,r}) = \infty$ (see Lemma \ref{lem: ball_bounds_in_probability}). In particular, for sufficiently large $n$ the condition $\min_{u \in \Csig[\Xbf]} \deg(u; \widetilde{G}[\Xbf]) \geq 10$ is satisfied, again with probability one. We may therefore apply Proposition \ref{prop: pointwise_mixing_time} and obtain
\begin{align*}
\liminf_{n \to \infty} \tau_{\infty}(\widetilde{G}_{n,r}) & \leq \liminf_{n \to \infty} \left\{\frac{2}{\Phi^2(\widetilde{G}_{n,r})} \log \left(\frac{1440}{s(\widetilde{G}_{n,r})}\right)\log \left(\frac{14}{s(\widetilde{G}_{n,r})}\right)  + 3 \log \left(\frac{14}{s(\widetilde{G}_{n,r})}\right) + 3 \right\}
\\
& \leq \frac{2^{27}\Lambda_{\sigma}^4\rho^2L^2 d^3}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{28800\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3 \log^2\left(\frac{280\Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + 3
\end{align*}
whereupon the theorem follows after an appropriate choice of universal constants $c_2, c_3, c_4 > 0$. In the rest of this section we proceed to proving Propositions \ref{prop: pointwise_mixing_time} and \ref{prop: local_spread_conductance}.

@@

We first generalize some notation from the main text. Let $A$ be the adjacency matrix of an undirected graph $G = (V,E)$, and $D$ the associated diagonal degree matrix. The lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf{} + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$ with elements $\pi_u = D_{uu}/\vol(V;G)$.  Denote the $t$-step probability distribution of this random walk, originating from a  vertex $v \in V$, as $q^{(t)}: V \times V \to [0,1]$, $q^{(t)}(v,u) = e_v \Wbf^t e_u$.

Consider the uniform distance\footnote{Note $d_{\textrm{unif}}$ is not formally a distance as it is not symmetric.} between the distributions $q_v^{(t)} := q^{(t)}(v,\cdot)$ and $\pi$, given by
\begin{equation*}
d_{\textrm{unif}}(q_v^{(t)},\pi) = \max_{u \in V} \set{\frac{\pi(u) - q_v^{(t)}(u)}{\pi(u)}}.
\end{equation*}
Our goal is to demonstrate that for a sufficiently large $t$, $d_{\textrm{unif}}(q_v^{(t)},\pi) \leq \frac{1}{4}$ for every $v \in V$ (see~\eqref{eqn: mixing_time} to recall the definition of $\tau_{\infty}(G)$). To do so, we introduce the \emph{total variation distance} between the distributions $q_v^{(t)}$ and $\pi$, given by
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sum_{u \in V} \abs{q_v^{(t)}(u) - \pi(u)}.
\end{equation*}
As we will see, an upper bound on the uniform distance can be obtained from an analogous upper bound on the total variation distance. First, however, we upper bound the total variation distance $||q_v^{(t)} - \pi||_{\mathrm{TV}}$ as a function of the local spread $s(G)$, the conductance $\Phi(G)$ and the number of steps $t$.
\begin{lemma}
	\label{lem: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1/4$,
	\begin{align}
	\norm{q_v^{(t + 3)} - \pi}_{\mathrm{TV}} \leq \max\set{as(G), \frac{1}{8} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} 
	\nonumber \\
	~~~~~~~~~~~~~~~~~~~~~ + \left(\frac{1}{1 - 2as(G)} +  \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t \label{eqn: tv_mixing_time}
	\end{align}
\end{lemma}

As mentioned, a bound on the total variation distance implies a corresponding bound on the uniform distance $d_{\mathrm{unif}}$, given by the following result:
\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{s(G)}{14}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}

\noindent Taking these Lemmas as given, we proceed that show that for
\begin{equation*}
\tau_1 = \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
\end{equation*}
the uniform distance $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$, which proves the claim of Proposition \ref{prop: pointwise_mixing_time}. Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$, so that
\begin{equation*}
\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}.
\end{equation*}
Recall that by assumption, $\min_{u \in V} \deg(u;G) \geq 10$. Moreover note that for $a \leq 1/4$, $2as(G) < 1/2 < 1 - 2as(G)$, since $s(G) \leq 9/10 < 1$. By Lemma \ref{lem: tv_mixing_time} we therefore obtain
\begin{equation*}
\norm{q_v^{(\tau_0 + 3)} - \pi}_{\mathrm{TV}} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{1}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
\end{equation*}
It is a well known fact (see e.g.~\cite{montenegro2002}) that if $\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \frac{1}{4}$, then for any $0 < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{\mathrm{TV}} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
\begin{equation*}
\norm{q_v^{(\tau_1)} - \pi}_{\mathrm{TV}} \leq \frac{s(G)}{14}
\end{equation*}
and so by Lemma \ref{lem: tv_to_uniform_distance}, $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. The proof of Proposition \ref{prop: pointwise_mixing_time} is therefore complete once we have proved Lemmas \ref{lem: tv_mixing_time} and \ref{lem: tv_to_uniform_distance}.

@@

We will find it useful to introduce the \emph{Lovasz-Simonovits curve}, defined for any $t \in \mathbb{N}$ and starting distribution $z$ to be $h_z^{(t)}: [0,1] \to [0,1]$,
\begin{equation*}
h_z^{(t)}(x) = \max_{w \in \mathcal{W}_x} \left\{ \sum_{u \in V} \left(q_z^{(t)}(u) - \pi(u)\right)w(u)\right\}.
\end{equation*}
The maximum in the preceding display is over the set of weight functions $\mathcal{W}_{x}$
\begin{align*}
\mathcal{W}_x = \left\{w: V \to [0,1] \Big\vert 0 \leq w(u) \leq 1~~\forall~u,~~\text{and}~~\sum_{u \in V} w(u) \pi(u) = x\right\}.
\end{align*}
The utility of the Lovasz-Simonovits curve is demonstrated by the following observations. First, note that
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} = \sup_{S \subseteq V} \abs{q_v^{(t)}(S) - \pi(S)}
\end{equation*}
where we use the standard notation $\pi(S) := \sum_{u \in S} \pi(u)$, and likewise for $q_v^{(t)}(S)$. Moreover, observe that for any $S \subseteq V$ and any integer $t \geq 0$,
\begin{equation}
\label{eqn: lovasz_simonovits_1}
\abs{q_v^{(t)}(S) - \pi(S)} \leq \max \set{h_v^{(t)}(\pi(S)), h_v^{(t)}(1 - \pi(S))}
\end{equation}
(To see this, consider the weight functions $w(u) = \1(u \in S)$ and $w'(u) = 1 - w(u)$.) Taking the maximum on both sides of \eqref{eqn: lovasz_simonovits_1}  over all $S \subseteq V$, we have
\begin{equation*}
\norm{q_v^{(t)} - \pi}_{\mathrm{TV}} \leq \max_{0 \leq x \leq 1} h_v^{(t)}(x).
\end{equation*}
Now, take $z = e_vW^3$, and trivially observe that $h_v^{(t+3)}(x) = h_z^{(t)}(x)$. To prove Lemma \ref{lem: tv_mixing_time}, it is therefore sufficient to show the desired upper bound -- the right hand side of~\eqref{eqn: tv_mixing_time}-- holds with respect to $h_z^{(t)}(x)$ for all $x \in [0,1]$, and all starting distributions $z = e_vW^3$.

Let $\mu = as(G)$, and note under the condition $a < 1/4$, $\mu < 1 - \mu$. To show the desired upper bound, we split the interval $[0,1]$ into the subinterval $[\mu, 1 - \mu]$ and the remainder $[0,\mu) \cup (1 - \mu, 1]$. Let $\ell_{\mu}(x)$ be the linear interpolator between $h_z^{(0)}(\mu)$ and $h_z^{(0)}(1 - \mu)$, 
\begin{equation*}
\ell_{\mu}(x) = \frac{1 - \mu - x}{1 - 2\mu} h_z^{(0)}(\mu) + \frac{x - \mu}{1 - 2\mu}h_z^{(0)}(1 - \mu).
\end{equation*}

The following technical Lemma collects the upper bounds which hold over $[\mu, 1 - \mu]$ and $[0,\mu) \cup (1 - \mu, 1]$, respectively:
\begin{lemma}
	\label{lem: mixing_all_sets}
	For any $\mu \leq x \leq 1 - \mu$ and any starting distribution $z = e_vW^3$,
	\begin{equation}
	\label{eqn: mixing_large_sets}
	h_z^{(t)}(x) \leq \ell_{\mu}(x) + \left(\frac{1}{1 - 2\mu} + \frac{1}{\mu}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t.
	\end{equation}
	For any $0 \leq x < \mu$ or $1 - \mu < x \leq 1$
	\begin{equation}
	\label{eqn: mixing_small_sets}
	h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{2^{t+3}} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
	\end{equation}
\end{lemma}
\noindent Taking Lemma \ref{lem: mixing_all_sets} as given, we have nearly completed our proof of Lemma \ref{lem: tv_mixing_time}. Note that for any $\mu \leq x \leq 1 - \mu$,
\begin{align*}
\ell_{\mu}(x) & \leq \max\set{h_z^{(t)}(\mu),h_z^{(t)}(1 - \mu)} \\
& \leq \max\set{as(G), \frac{1}{2^{t + 3}} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} }
\end{align*}
with the latter inequality following from \eqref{eqn: mixing_small_sets}. Therefore, for any $x \in [0,1]$,
\begin{equation*}
h_z^{(t)}(x) \leq \max\set{as(G), \frac{1}{8} + \frac{9a}{20} + \frac{1}{2 \min_{u \in V}\deg(u;G)} } + \left(\frac{1}{1 - 2as(G)} + \frac{1}{as(G)}\right)\left(1 - \frac{\Phi^2(G)}{2}\right)^t
\end{equation*}
which is exactly the claimed result of Lemma \ref{lem: tv_mixing_time}.

@@

We prove~\eqref{eqn: local_spread} immediately before turning our attention to~\eqref{eqn: conductance}, which will require the bulk of our attention. First, however, we introduce some notation. Recall that $\widetilde{G}_{n,r} = G_{n,r}\bigl[\Csig[\Xbf]\bigr]$ denotes the subgraph induced by $\Csig[\Xbf]$. For $S \subseteq \Csig[\Xbf]$ and $u \in \Csig[\Xbf]$, we will write
\begin{equation*}
\widetilde{\cut}_{n,r}(S) := \cut(S; \widetilde{G}_{n,r}), \quad \widetilde{\deg}_{n,r}(u) := \deg(u; \widetilde{G}_{n,r}), \quad 
\widetilde{\vol}_{n,r}(S) := \vol(S; \widetilde{G}_{n,r})
\end{equation*}
and let $\widetilde{\pi}_{n,r}(u) = \widetilde{\deg}_{n,r}(u)/\widetilde{\vol}_{n,r}(\Csig[\Xbf])$ be the stationary distribution of the lazy random walk over $\widetilde{G}_{n,r}$. We also let $\widetilde{\Phi}_{n,r}(S) := \Phi(S; \widetilde{G}_{n,r})$ denote the normalized cut functional over $\widetilde{G}_{n,r}$.
